---
title: "CMSC320 Final Project"
author: David Wang, Steve Jettoo
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ingesting and Tidying Data

```{r start}
library(tidyverse)
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(crypto)

crypto_hist <- crypto_history(coin = "bitcoin", limit = NULL, start_date = "20170427", end_date ="20190228")

colnames(crypto_hist)[colnames(crypto_hist)=="date"] <- "Date"
 
crypto_hist
```



```{r excel_data}
library(readxl)

CMSC320_Crypto_Data <- read_excel("CMSC320 Crypto Data.xlsx")
excel_data <- data.frame(CMSC320_Crypto_Data) %>%
   mutate(Date = as.Date(Date, format="%Y-%m-%d"))
excel_data

```

Next, read the data from "CMSC320 Crypto Data.xlsx". This dataset is an excel spreadsheet. The data within the excel spreadsheet is data from various sources. 

These are the attributes within the dataset and a short explanation of what they are:
TxPerDay: Number of Transaction over a 24 hour period. 
AvgFee: Daily aggregate fee added to transaction (transaction cost)
Hashrate: Daily relative measure of network's processing power in trillions of terahashes per second
Diff: Daily mathematical value of how hard it is to find a valid hash for a block
ResElec: Daily residential Electricity Price in cents/kWh
CommElec: Daily commercial Electricity Price in cents/kWh
FFR: Federal Funds Rate or the interest rate at which banks lend reserve balances to other banks overnight
Trending: Daily measure of interest in bitcoin over time (from Google trends)
Infl: Inflation Rate; used as a macroeconomic indicator on the impact of price changes on Bitcoin value
TT: A unique identifier for every observation

To import an excel spreadsheet onto R, one must first have the package library readxl installed. The package readxl allows R to easily access the data within Excel. We will use the function 'read_excel' to read the xlsx file. Once we have the excel file in R, we will transform it so that it becomes a data frame. This is so we are able to manipulate the data within the excel sheet. We will use the function 'data.frame' to transform the xlsx file into a data frame.

Notice that the type for the attribute Date is different than the one in the first dataset. It is essential for the Date attributes from the two datasets to be the same type in order for them to merge with each other. The 'mutate' function allows us to create new attributes based on the result of a given expression. We will use 'mutate' to transform Date into a type <date>. The expression we will use in the 'mutate' function will be the 'as.Date' function, which will convert character data to dates. The 'as.Date' function will allow us to convert the Date attribute into type date. 

Once the Date attribute has been converted into type date, we now have a data set that can be merged with our previous dataset.

```{r merge_data}
crypto_merge <- crypto_hist %>%
  left_join(excel_data, by = "Date") 
crypto_merge
```

We will now merge the two datasets together using their Date attribute. To do this, we will be using the 'join' operation. The 'join' operation allows us to merge the rows of two data sets forming one data set, given a matching attribute from the two data sets. The join operation that we will be mainly using is the 'left_join' operation. 'left_join' will merge the two data sets together; however, all observations on the left hand side or the first data set are retained.

We will merge the data set excel_data onto the data set crypto_hist using the common attribute "Date" with the function 'left_join'. This will create one data set that contains the data from both data sets. We will call this new data set, crypto_merge. 


```{r DJW}
# Gets stock data from yahoo
# install.packages("quantmod")

library(quantmod)

# Scrape Data from yahoo finance
getSymbols("^DJI", from = "2017-4-27", to = "2019-2-28")

# Set it as a data frame
DJI_data <- data.frame(DJI) 
DJI_data
```

We will now injest stock market data. To do this we require the quantmod package. The quantmod package allows us to easily access stock data in R. For this project, we want stock information about the Dow Jones Industrial Average (^DJI). To get this information, we will use the function 'getSymbols' from the quantmod package. The 'getSymbols' function allows us to load and manage 'Symbols' in a specified enviroment. We will use it to obtain stock data for the Dow Jones Industrial Average. The function 'getSymbols' should create an xts object with the name of the stock company when loading the stock data of a specific company. In this case the xts object with the stock data should be named "DJI". Tranform the xts object into a data frame using data.frame. We will call this DJI_data.

Further Reading for the quantmod package: https://cran.r-project.org/web/packages/quantmod/quantmod.pdf

```{r rename_DJI}
# Date is set as a row name. We need to change this so that Date becomes an attribute. This will be neccessary if we want to merge the stock data with the bitcoin data

# Add a column name to the dates as well as change the names for the rest of the attributes
DJI_data <- cbind(rownames(DJI_data), DJI_data)

# Removes the row name
rownames(DJI_data) <- NULL
colnames(DJI_data) <- c("Date","DJI_Open","DJI_High", "DJI_Low", "DJI_Close", "DJI_Volume", "DJI_Adjusted")

DJI_data 
```


When viewing the stock data for DJI, notice that the dates are row names instead of values. We need to tranform the dates for a row name to a value. This is so we can merge this data set into our crypto_merge data set. To do this we will first use the function 'cbind' to create a new attribute using the dates that are row names. We then will remove the row names by setting them to be NULL. Finally, we will rename the column names using the function 'colnames'. 

Here are the attributes and an explanation of what each attribute is for this data set: 
DJI_Open: The opening price of a stock of the Dow Jones Industrial Average
DJI_Close: The closing price of a stock of the Dow Jones Industrial Average
DJI_Volume: The number of Dow Jones Industrial Average shares that changed hands during a given day
DJI_Adjusted: The adjusted closing price of a stock of the Dow Jones Industrial Average taking into account any corporate actions


```{r NVDA}
getSymbols("NVDA", from = "2017-4-27", to = "2019-2-28")

NVDA_data <- data.frame(NVDA) 
NVDA_data

```

We will repeat the same process to ingest and scrape Nvidia stock data. We will use the 'getSymbols' function to get Nvidia stock data, and then we will transform it into a data frame. We will call this NVDA_data. 

```{r rename_NVDA}
NVDA_data <- cbind(rownames(NVDA_data), NVDA_data)

rownames(NVDA_data) <- NULL
colnames(NVDA_data) <- c("Date","NVDA_Open","NVDA_High", "NVDA_Low", "NVDA_Close", "NVDA_Volume", "NVDA_Adjusted")
NVDA_data
```

We will now fix the data so that dates are in an attribute. We will the same process as before. 

Here are the attributes and an explanation of what each attribute is for this data set:
NVDA_Open: The opening price of a stock of Nvidia
NVDA_Close: The closing price of a stock of the Nvidia
NVDA_Volume: The number of Nvidia shares that changed hands during a given day
NVDA_Adjusted: The adjusted closing price of a stock of the Nvidia taking into account any corporate actions

```{r merge_stocks}
stock_data <- NVDA_data %>%
  left_join(DJI_data, by="Date") %>%
  mutate(Date = as.Date(Date, format="%Y-%m-%d"))
stock_data
```

We will now merge the the DJI stock data set with the NVDA stock data set using the 'left_join' function with the common attribute "Date". The two data sets will be merged into one data set that we will call stock_data.

Notice that the data type for Date in stock_data is not the same as the data type for Date in crypto_merge. We need to change the data type of "Date"" in stock_data in order to properly merge stock_data and crypto_merge together. We will use the 'as.Date' function to change the "Date" data type from a <fctr> data type to a <date> data type

```{r merge_stock}
crypto_data <- cryto_merge %>%
  left_join(stock_data, by="Date")
crypto_data
```

We will now merge the data set crypto_data with the data set stock_data using the 'left_join' function using the common attribute "Date". This will merge all the data that we ingested into one data set. We will call this crypto_data.

## Missing Data: Imputation

```{r indicate_missing}
cryp_up <- crypto_data %>%
  mutate(stock_missing = is.na(NVDA_Open))
  
crypto_data <- cryp_up
crypto_data
```

Looking at the data set, crypto_data, we can see that the stock attributes for NVDA and DJI have missing data. This is because the stock market is not operating during certain days. 

In regards to missing data,it is important to know why the data was missing in order to determine how to approach missing data. There are two different apporaches regarding missing data. One could either remove the observations with missing data or one could impute the missing data. This usually depends on whether the data was missing at random or systematically. 

But before that, let us first add an additional indicator variable stating if the stock data is missing or not. We will use the function 'is.na' to determine whether or not the stock is missing or not. We will also use the 'mutate' function to create the new indicator variable. Given that both NVDA and DJI will have missing values on the same day, we will only need one indicator variable to state if the value is missing or not. 


```{r missing}
crypto_data$NVDA_Open[is.na(crypto_data$NVDA_Open)] <- mean(crypto_data$NVDA_Open, na.rm=TRUE)
crypto_data$NVDA_High[is.na(crypto_data$NVDA_High)] <- mean(crypto_data$NVDA_High, na.rm=TRUE)
crypto_data$NVDA_Low[is.na(crypto_data$NVDA_Low)] <- mean(crypto_data$NVDA_Low, na.rm=TRUE)
crypto_data$NVDA_Close[is.na(crypto_data$NVDA_Close)] <- mean(crypto_data$NVDA_Close, na.rm=TRUE)
crypto_data$NVDA_Volume[is.na(crypto_data$NVDA_Volume)] <- mean(crypto_data$NVDA_Volume, na.rm=TRUE)
crypto_data$NVDA_Adjusted[is.na(crypto_data$NVDA_Adjusted)] <- mean(crypto_data$NVDA_Adjusted, na.rm=TRUE)

crypto_data$DJI_Open[is.na(crypto_data$DJI_Open)] <- mean(crypto_data$DJI_Open, na.rm=TRUE)
crypto_data$DJI_High[is.na(crypto_data$DJI_High)] <- mean(crypto_data$DJI_High, na.rm=TRUE)
crypto_data$DJI_Low[is.na(crypto_data$DJI_Low)] <- mean(crypto_data$DJI_Low, na.rm=TRUE)
crypto_data$DJI_Close[is.na(crypto_data$DJI_Close)] <- mean(crypto_data$DJI_Close, na.rm=TRUE)
crypto_data$DJI_Volume[is.na(crypto_data$DJI_Volume)] <- mean(crypto_data$DJI_Volume, na.rm=TRUE)
crypto_data$DJI_Adjusted[is.na(crypto_data$DJI_Adjusted)] <- mean(crypto_data$DJI_Adjusted, na.rm=TRUE)

crypto_data
```

Regarding the missing stock data, we will use imputation. Since the stock data is all numeric, we will simply replace the missing values with the mean of the non-missing values. The result of this method will have two effects. One, it will retain the central tendancy of the attributes as the mean will not change. Two, the spread of the imputed attributes will have smaller spreads compared to the spread if we have chosen to ignore the missing values. A smaller spread may be hazardous as it causes over-confident inferences during analysis. It is important to recognize this when deciding how to handle missing data.

To replace the missing values with mean of the non-missing values, we will use the function 'mean' and we will set it so it will remove the missing data by setting na.rm = TRUE. We will do this for all stock variables. Additionally, we will check if a value is missing using the 'is.na' function, which will return TRUE if a value is missing and FALSE otherwise.

We have now filled in the missing data for our stock variables. Now the next step is to analyze the data to determine what affects the price of Bitcoin.

## Regression

```{r plot}
crypto_data %>%
  ggplot(aes(x=Date, y=close)) +
    geom_point() +
    geom_line() + 
    labs(title="Bitcoin Value Over Time",
         x = "Time",
         y = "Value of Bitcoin")

```

In order to analyze what determines the price of Bitcoin, we will use a logarithmic regression. The reason we will be performing a logarithmic regression rather than a linear regression is due the exponential nature of the price of Bitcoin. This can be seen in the plot, which was created using the function 'ggplot' with the x-axis being Time and the y-axis being the Value of Bitcoin.

We will run a regression with the closing price of Bitcoin, close, as our dependent variable. This will be used to represent the value of Bitcoin. This will be logged due to its exponential nature. To log a variable, use the 'log' function. We will also log several of the independent variables due to their exponential nature. 

From the data set, crypto_data, we selected various variables as we believe that they may have an affect on the value of Bitcoin. Here is our intuition:

Volume: Indicates the amount of money people are putting into the Bitcoin system. The larger the volume the more confident people are with the Bitcoin, which would increase its value.

Spread: Indicates how speculation could have an effect on the value of Bitcoin.

TxPerDay: The more transactions individuals are making through Bitcoin, the more accepting people are with Bitcoin, which would increase its value. This will be logged.

AvgFee: A higher fee would indicate a higher demand for Bitcoin, which may increase its value. This will be logged.

Hashrate: A higher hashrate indicates that more people are involved in the Bitcoin system as well as it being more secure. This will be logged.

Diff: The harder it is to mine a Bitcoin the less return on investment one would get to mine it possibly decreasing its value. This will be logged.

ResElec/CommElec: Input costs to mine Bitcoin. A higher price in electricity may decrease the value of Bitcoin

FFR: Higher interest rates means that miners can not borrow much money to pay for hardware to mine Bitcoin, which may decrease the value of Bitcoin.

Trending: The popularity of Bitcoin may affect its price. This will be logged.

Infl: Chosen as a macroeconomic indicator of the impact of price changes on Bitcoin value

DJI_Adjusted: Instrument for general health of the US economy. This will be logged.

NVDA_Adjusted: Chosen as an instrument for interest in Bitcoin mining. This will be logged.



```{r regression}
price <- lm(log(close) ~ log(volume) + spread + log(TxPerDay) + log(AvgFee) + log(Hashrate) + log(Diff) + ResElec + CommElec + FFR + log(Trending) + Infl + log(NVDA_Adjusted) + log(DJI_Adjusted), data = crypto_data)

broom::tidy(price)
```

We will now perform a logarithmic regression based on the dependent and indenpendet variables that we have chosen. We will use the function  

On average, holding everything else constant, a one percent increase in volume will increase the price of Bitcoin by 0.392696910479%.

On average, 


## Regression Analysis

```{r fitted}
check <- broom::augment(price) 
check %>%
  ggplot(aes(x=.fitted, y=.resid)) +
    geom_point() +
    geom_smooth() +
    labs(title="Fitted vs. Residuals",
         x = "fitted",
         y = "residuals")
```


## Random Forest

```{r forest}
set.seed(1234)
new_cryp <- na.omit(crypto_data)
train_indices <- sample(nrow(new_cryp), nrow(new_cryp)/2)
train_set <- new_cryp[train_indices,]
test_set <- new_cryp[-train_indices,]

library(randomForest)

cryp_rf <- randomForest(close ~ volume + spread + TxPerDay + AvgFee + Hashrate + Diff + ResElec + CommElec + FFR + Trending + Infl + NVDA_Adjusted + DJI_Adjusted, importance=TRUE, mtry=10, data=train_set)

plot(cryp_rf)
```

```{r interpret}
variable_importance <- importance(cryp_rf)
knitr::kable((round(variable_importance, digits=2)))
```

```{r plot}
varImpPlot(cryp_rf)
```

## Leaflet Visualization
```{r Local Volume Data}
# need readxl library
library(readxl)

# read excel on local volume data in repository
volume_data <- read_excel("localvolumedata.xlsx")

# Our Data came out a bit wonky, so lets clean it up.
volume_data_df <- volume_data %>%
  as_data_frame()

# Regexpression
countryrexp <- "([A-Z]+[a-z]*)(\\s)*([A-Z]{1}[a-z]+)*.*"

# Clean up Country Data Values
# Change the name of some countries to make joining data easier
localvolumedata <- volume_data_df %>%
  mutate(Country = str_extract(Country,countryrexp))

# Change the name of some countries to make joining data easier
localvolumedata$Country[localvolumedata$Country == "USA"] <- "United States of America"
localvolumedata$Country[localvolumedata$Country == "Hong Kong"] <- "Hong Kong S.A.R."
localvolumedata$Country[localvolumedata$Country == "Tanzania"] <- "United Republic of Tanzania"

```

```{r Leaflet}
library(rgeos)
library(rworldmap)
library(leaflet)
library(rgdal)
library(jsonlite)
library(xts)

# get world map
# note must also install rworldextra

wmap <- getMap(resolution= "high")

# get centroids
centroid <- gCentroid(wmap, byid=TRUE)

# get a data.frame with centroids
mapcen <- as.data.frame(centroid)

mapcen <- cbind(rownames(mapcen), mapcen)

rownames(mapcen) <- NULL
colnames(mapcen) <- c("Country", "Lng", "Lat")

maploc <- mapcen %>%
  mutate(Country = as.character(Country))


combined_map <- localvolumedata %>%
  inner_join(maploc, by = "Country")

combined_map <- combined_map %>%
  mutate(pcolor = case_when(
    Percentage < .001 ~ rgb(241/255,196/255,15/255),
    Percentage < .01  ~ rgb(243/255,156/255,18/255),
    Percentage < .1   ~ rgb(230/255,126/255,34/255),
    Percentage < .2   ~ rgb(211/255,84/255,0/255),
    Percentage >= .2   ~ rgb(135/255,54/255,0/255)
    ))

combined_map

ico <- awesomeIcons(
  icon = "bitcoin",
  library = "glyphicon",
  markerColor = "white",
  iconColor = combined_map$pcolor
)

colorspal <- c(rgb(241/255,196/255,15/255),rgb(243/255,156/255,18/255),rgb(230/255,126/255,34/255),rgb(211/255,84/255,0/255),rgb(135/255,54/255,0/255))


world_map <- leaflet(combined_map) %>%
  addTiles() %>%
  addAwesomeMarkers(~Lng,~Lat, icon = ~ico, popup = paste("<b>Country: </b> ", combined_map$Country, "<br>","<b>Continent: </b> ", combined_map$Continent, "<br>","<b>Native Currency: </b> ", combined_map$`Native Currency`, "<br>","<b>FX Rate: </b> ", combined_map$`FX Rate`, "<br>","<b>USD Value: </b> ", combined_map$`USD Value`, "<br>", "<b>Percentage: </b> ", combined_map$Percentage, "<br>"), clusterOptions = markerClusterOptions(), label = ~Country) %>%
  addLegend(labels = ~c("<b>x < 0.1%</b>","<b>x < 1%</b>","<b>x < 10%</b>","<b>x < 20%</b>","<b>x >= 20%</b>"), colors = ~colorspal, opacity = 1, title= "Volume Percentage")
  
world_map
```

